---
title: "Group 2, Project 2 WIP"
author: "Nathan Koh"
date: "2024-11-17"
output: html_document
---


loading csv
```{r}

data<-read.csv("kc_house_data.csv", header=TRUE)
#removing NAs
data[!complete.cases(data),]
```

From Kaggle, variable definitions:

*id* - Unique ID for each home sold
*date* - Date of the home sale
*price* - Price of each home sold
*bedrooms* - Number of bedrooms
*bathrooms* - Number of bathrooms, where .5 accounts for a room with a toilet but no shower
*sqft_living* - Square footage of the apartments interior living space
*sqft_lot* - Square footage of the land space
*floors* - Number of floors
*waterfront* - A dummy variable for whether the apartment was overlooking the waterfront or not

Categorical
-------------------
*view* - An index from 0 to 4 of how good the view of the property was
*condition* - An index from 1 to 5 on the condition of the apartment,
*grade* - An index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design.
-------------------
*sqft_above* - The square footage of the interior housing space that is above ground level
*sqft_basement* - The square footage of the interior housing space that is below ground level


Build Details + Geography
--------------------------
*yr_built* - The year the house was initially built
*yr_renovated* - The year of the house’s last renovation
*zipcode* - What zipcode area the house is in
*lat* - Latitude
*long* - Longitude
*sqft_living15* - The square footage of interior housing living space for the nearest 15 neighbors (create viz vs price)
*sqft_lot15* - The square footage of the land lots of the nearest 15 neighbors (create viz vs price)

TO DO: Find another error by Monday, 11/25. Do #6
Observations WIP:

1) The Bathroom column indicates 0.5 represents a bathroom with a toilet but no shower. Data shows 0.25 increments.

2) Comparing price to other columns may not make sense

3) Bedroom, Bathroom, and the Price comparisons on certain homes do not make sense

4) Floors indicate a 0.5th floor but is not described as to what it represents



Removing the following vars: zipcode,lat, long. Location and geography not as important
Re-adding yr_built, yr_renovated to consider for cost based on the data
```{r}
library(dplyr)
library(tidyverse)

data<- data %>% select(-zipcode,-lat,-long)

data
```

Checking for duplicate entries
```{r}
#unique IDs have duplicates from result. ~200 homes 
distinct.sales<-data %>% distinct(id, .keep_all= TRUE)
distinct.sales
```

Checking for details when bedroom, bathroom = 0
```{r}
#comparison 
data %>% 
  group_by(bedrooms,bathrooms) %>% filter(bedrooms==0,bathrooms==0) #maybe remove the 0 b/b's since points are in single digits
```

Several columns in the 0 b/b category do not make sense or are entry errors - price, category, grade. Yr_built also does not make sense in context



```{r}
#creating bathroom df of 0.25 and 0.75

distinct.sales[round(distinct.sales$bathrooms %% 1, 2) ==0.25 | round(distinct.sales$bathrooms %% 1, 2) == 0.75,]
```

7229 homes with a partial bathroom ending in .25 or .75. Values ending in 0.5 means there is a toilet but no shower. What do numbers ending in 0.25 / 0.75 mean in context?

ID 2524049179 is priced at $2m with 3 bedrooms / 2.75 bathrooms and a sqft lot of 44867 which is significantly smaller than a majority of the homes listed. If it's a typo, consider deleting one zero, one four. (Open to discussing alternative solutions)

```{r}
#sorting by sqft in descending order
distinct.sales[order(-distinct.sales$sqft_lot),]
```
```{r}
#visualization on price against interior size in sqft
library(ggplot2)
ggplot(data, aes(x=sqft_living, y=price))+
  geom_point()+
  labs(x="sqft_living", y="price", title="price - sqft living comparison")

```

6) Provide data visualizations that explore the what characteristics are associated with good quality homes whose condition is greater than 7.
```{r}
#filter df so that it only shows homes of condition >3, grade >7
good.df<- data %>% 
  group_by(condition,grade) %>% filter(condition>3,grade>7)

good.df #2652 homes
```

test visualization of grade / condition against price
```{r}
#number of homes per condition tier
ggplot(data,aes(x=condition))+
  geom_text(aes(label=..count..),stat='count', position = position_dodge(0.9),vjust=-0.2)+
  geom_histogram(binwidth=1,fill="blue",color="orange")
```

```{r}
#number of homes per grade tier
ggplot(data,aes(x=grade))+
  geom_text(aes(label=..count..),stat='count', position = position_dodge(0.9),vjust=-0.2)+
  geom_histogram(binwidth=1,fill="blue",color="orange")

#observations: Most homes are graded 7 + 8
```

```{r}
#filtering to good.df

ggplot(good.df,aes(x=condition))+
  geom_text(aes(label=..count..),stat='count', position = position_dodge(0.9),vjust=-0.2)+
  geom_histogram(binwidth=1,fill="blue",color="orange")
```

```{r}
#5 number summary of prices where condition >3, grade >7
summary(good.df$price)
```

Biggest outlier is the home for $7.7m. Median price is 682000, mean price is 793543, minimum price is 149900

#scatterplot of condition and grade to sqft
```{r}
library(dplyr)
library(tidyverse)

ggplot(good.df,aes(x=bedrooms))+
  geom_text(aes(label=..count..),stat='count', position = position_dodge(0.9),vjust=-0.2)+
  geom_histogram(binwidth=1,fill="blue",color="orange")

good.df2<-good.df %>% 
  mutate(bedroom=ifelse(bedrooms<=4,"Less 4 Bedrooms","Greater 4 bedrooms"))

twoway<-table(good.df$grade,good.df2$bedroom)
twoway

prop.table(twoway,1)
```
There are more homes that have less than 4 bedrooms but two-way does not show much based on comparison.
```{r}
twoway.condition<-table(good.df$condition,good.df2$bedroom)
twoway.condition
```
Likewise, there are more homes that consist of less than 4 bedrooms whose condition is either a 4 or 5. Bedrooms might not be a factor to determining the condition or grade of a house. (could do stacked plot if needed)



```{r}
good.df$condition<-factor(good.df$condition)

#condition to sqft_lot to look for observations
ggplot(good.df,aes(x=condition,y=sqft_lot,fill=condition))+
  geom_boxplot()+ylim(0,80000)
```
```{r}
ggplot(good.df,aes(x=condition,y=price,fill=condition))+
  geom_boxplot()
```

Observations:
1) The IQR's are slightly different for conditions 4 and 5, but the median sqft_lot appears to be closely similar to each other. (Should we keep outliers in, remove, or have both visuals)

2) Comparing condition and price, the median price for homes with condition 4 are less than that of condition 5. Makes sense in this context as prices should increase as the quality improves. (There may be less outliers in price for 5 than 4)
```{r}
#do the same with grade
good.df$grade <-factor(good.df$grade)

ggplot(good.df,aes(x=grade,y=sqft_lot,fill=grade))+
  geom_boxplot()+ylim(0,80000)

```
```{r}
#homes with basements

basement <- good.df %>% 
  group_by(sqft_basement) %>% 
  filter(sqft_basement!=0)

basement<- basement %>%  select(-zipcode,-lat,-long)


```

There are 1536 of 21613 homes that have basements and fulfill both condition and grade requirements.

Note: sqft_basement is noise and does not affect price as thought to be. Variables to consider are View and Waterfront.

\
```{r}
library(dplyr)
library(tidyverse)
good.df$waterfront<-factor(good.df$waterfront)
waterfront.df<-good.df %>% 
  mutate(wfront=ifelse(waterfront==1,"Waterfront","No Waterfront"))


ggplot(good.df,aes(x=view,y=price,colour =waterfront))+
  geom_point()
```
Visualization comparing view to price shows to have a slight average price increase but shows a clear outlier for the home with view of 3 and condition / grade requirement met. ID is 6762700020. There is also an outlier when view =4 for the same reason. ID is 7558700030.

Most of the homes with waterfronts are related to when it has a view of 4 (some in 3). But a majority of the homes in this dataset do not have waterfronts. We can consider that waterfronts may not affect price.

```{r}
good.df %>% 
  group_by(view) %>% 
  summarize(mean.price=mean(price))
```

```{r}
good.df %>% 
  group_by(waterfront) %>% 
  summarize(mean.price=mean(price))
```

Creating a boxplot with both observations:
```{r}
good.df$view<-factor(good.df$view)
ggplot(good.df,aes(x=waterfront,y=price,fill=view))+
  geom_boxplot()
```
Observations from Boxplot:
1) The median line increases based on the view. Safe to say that the overall price of the house increases due to similar conditions
2) The homes with waterfronts have a non-zero view rating
3) 0 waterfront homes with 0 view have the most outliers. 
4) The 3 view boxplot with waterfront = 1 has a lesser median price than that of a home with view =2. This can be assumed to be due to sample size differences. (~50 datapoints for homes)
5) Outlier points due to data entry error: Waterfront=0, View =3, price is 7.7m


WIP:
1) yr_renovated to price comparison