---
title: "Project 2"
author: "James Caldwell"
output: pdf_document
---

```{r,echo=FALSE,results='hide', message=FALSE}
rm(list = ls())
library(dplyr)
library(MASS)
library(ggplot2)
library(MASS) ##to use boxcox function
library(GGally)
library(faraway)
# ctrl alt i is shortcut for adding a new r chunk
# Shortcut is Ctrl + Shift + M for %>% symbols

```

```{r}
# Import Data
Data <- read.csv("kc_house_data.csv",header = TRUE)
zipcode <- read.csv("zipcodeData.csv")
# Add new variable, Region (based on zip code
Data$region <- c()
for (i in 1:length(Data$zipcode)){
  regionMatch <- zipcode[zipcode$x == Data$zipcode[i],]$region
  Data$region[i]<- regionMatch 
}
# Remove errors as discussed in section 3
cleanData <- Data[(Data$bedrooms!=0),]
cleanData <- cleanData[(cleanData$bathrooms!=0),]
cleanData <- cleanData[(cleanData$bedrooms<30),]
cleanData <- cleanData[ , -which(names(cleanData) %in% c("date","sqft_above","sqft_basement","zipcode","lat","long"))]

set.seed(6021)
sample.data<-sample.int(nrow(cleanData), floor(.50*nrow(Data)), replace = F)
train<-cleanData[sample.data, ]
test<-cleanData[-sample.data, ]

# Add training/test data

regnull <- lm(price~1, data=train) # Intercept only model
regfull <- lm(price~., data=train) # Model with all predictors

MLR_forward <- step(regnull, scope=list(lower=regnull, upper=regfull), direction="forward")

```

```{r}
# Extract AIC values
aic_values <- MLR_forward$anova$AIC
# Create a data frame for ggplot
aic_data <- data.frame(Step = seq_along(aic_values),AIC = aic_values)

ggplot(aic_data, aes(x = Step, y = AIC)) +
  geom_line(color = "blue", size = 1) +      # Line connecting points
  geom_point(color = "blue", size = 2) +     # Points
  labs(
    title = "AIC at Each Step of Forward Selection",
    x = "Step",
    y = "AIC" ) +
  geom_vline(xintercept = 6) +
  geom_hline(yintercept = 265000)
```
(before using Sarah's region)
Step 6: 
price ~ sqft_living + lat + view + grade + yr_built


with zip code regions:
price ~ sqft_living + region + waterfront + grade + yr_built

Check residual plots:
```{r}
result<-lm(price ~ sqft_living + region + waterfront + grade + yr_built, data=train)
par(mfrow = c(2, 2))
plot(result)
```
```{r}
MASS::boxcox(result)
```


```{r}
##transform y 
ystar<-log(train$price)
train<-data.frame(train,ystar)

result2<-lm(ystar ~ sqft_living + region + waterfront + grade + yr_built, data=train)
par(mfrow = c(2, 2))
plot(result2)
```
Investigating high leverage points:
```{r}
hii<-lm.influence(result2)$hat ##leverages
ext.student<-rstudent(result2) ##ext studentized res
n<-nrow(cleanData)
p<-5

length(hii[hii>2*p/n])

```
~6000 data points out of 22000 are considered high leverage.

Outliers:
```{r}
length(ext.student[abs(ext.student)>3])
```
89 data points could be considered outliers.

Influential observations:
```{r}
DFFITS<-dffits(result)
length(DFFITS[abs(DFFITS)>2*sqrt(p/n)])
```
Using the DEFITS method, there are ~1200 data points that could be influential. 


The final model is:
```{r}
print(result2$coefficients)
```

Make a model with all the predictors and bad predictors to compare with our model:
```{r}
model_full<-lm(ystar ~ sqft_living + region + waterfront + grade + yr_built + 
    view + bedrooms + bathrooms + condition + sqft_lot15 + sqft_living15 + 
    id + floors + sqft_lot, data=train)
model_bad<-lm(ystar ~ floors + sqft_lot, data=train)

test$log_price_predictions_full = predict(model_full,newdata = test)
test$price_predictions_full = exp(test$log_price_predictions_full)

test$log_price_predictions_bad = predict(model_bad,newdata = test)
test$price_predictions_bad = exp(test$log_price_predictions_bad)
```



## Prediction on test data and evaluation of model accuracy

```{r}
test$log_price_predictions = predict(result2,newdata = test)
test$price_predictions = exp(test$log_price_predictions)

test$log_price_predictions_bad = predict(model_bad,newdata = test)
test$price_predictions_bad = exp(test$log_price_predictions_bad)

test$log_price_predictions_full = predict(model_full,newdata = test)
test$price_predictions_full = exp(test$log_price_predictions_full)

# Print the first 25 rows
test_price_only <- test %>%
  dplyr::select(price, price_predictions,sqft_living,grade,yr_built)

head(test_price_only,25)

library(tidyr)

# Create a long format dataframe with a combined price and price predictions, with 
  # "Type" as a new column denoting which one each entry is. This way, we can plot
  # The two on the same plot to visualize how the model is predicting vs the actuals
test_long <- test %>%
  pivot_longer(cols = c(price, price_predictions,price_predictions_bad,price_predictions_full),
               names_to = "type",
               values_to = "predicted_and_actual")

# Create the plot
ggplot(test_long, aes(x = sqft_living, y = predicted_and_actual, color = type)) +
  geom_point(size = 1, alpha = 1) +
  labs(title = "Actual vs Predicted Prices",
       x = "Square Footage",
       y = "Price") +
      xlim(0, 8000) +  # Set x-axis limits
      ylim(0, 7000000)   # Set y-axis limits

# This plot skips displaying one datapoint, which is the very expensive house of 1.25M Sqft. Zooming in beyond this point helps see the smaller details better.
```
RMSE
```{r}
library(Metrics)

# Calculate RMSE
rmse_small <- rmse(test$price, test$price_predictions)
cat('Ideal model:',rmse_small)
rmse_bad <- rmse(test$price, test$price_predictions_bad)
cat('\nBad model:',rmse_bad)
rmse_full <- rmse(test$price, test$price_predictions_full)
cat('\nfull model:',rmse_full)
```

```{r}
test$residual_percent <- (abs(test$price-test$price_predictions_full)/test$price) * 100

ggplot(test, aes(x = price, y =residual_percent)) +
  geom_point() +
  labs(title = "Percent Accuracy of predicted vs actual",
       x = "Price",
       y = "Percent Accuracy")

summary(test$residual_percent)
```
```{r}
# test_filtered <- test %>%
#   filter(price > 300000) %>%
#   mutate(residual_percent_filtered = residual_percent)
# 
# ggplot(test_filtered, aes(x = price, y =residual)) +
#   geom_point()
# 
# summary(test_filtered$residual_percent)
```

