---
title: "Project 2"
author: "James Caldwell"
output: pdf_document
---

```{r,echo=FALSE,results='hide', message=FALSE}
rm(list = ls())
library(dplyr)
library(MASS)
library(ggplot2)
library(MASS) ##to use boxcox function
library(GGally)
library(faraway)
library(leaps)
library(car)
# ctrl alt i is shortcut for adding a new r chunk
# Shortcut is Ctrl + Shift + M for %>% symbols

```

Remove sqft above/below, date, region instead of zip/lat/long


```{r}
Data<-read.csv("kc_house_data.csv", sep=",", header=TRUE)
zipcode <- read.csv("zipcodeData.csv")
Data$region <- c()
for (i in 1:length(Data$zipcode)){
  regionMatch <- zipcode[zipcode$x == Data$zipcode[i],]$region
  Data$region[i]<- regionMatch 
}


# Add training/test data

regnull <- lm(price~1, data=Data) # Intercept only model
regfull <- lm(price~., data=Data) # Model with all predictors

MLR_forward <- step(regnull, scope=list(lower=regnull, upper=regfull), direction="forward")

```

```{r}
# Extract AIC values
aic_values <- MLR_forward$anova$AIC
# Create a data frame for ggplot
aic_data <- data.frame(Step = seq_along(aic_values),AIC = aic_values)

ggplot(aic_data, aes(x = Step, y = AIC)) +
  geom_line(color = "blue", size = 1) +      # Line connecting points
  geom_point(color = "blue", size = 2) +     # Points
  labs(
    title = "AIC at Each Step of Forward Selection",
    x = "Step",
    y = "AIC" ) +
  geom_vline(xintercept = 6) +
  geom_hline(yintercept = 529000)
```
(before using Sarah's region)
Step 6: 
price ~ sqft_living + lat + view + grade + yr_built

with zip code regions:
price ~ sqft_living + region + waterfront + grade + yr_built

Check residual plots:
```{r}
result<-lm(price ~ sqft_living + region + waterfront + grade + yr_built, data=Data)
par(mfrow = c(2, 2))
plot(result)
```


```{r}
##transform y 
ystar<-log(Data$price)
Data<-data.frame(Data,ystar)

result2<-lm(ystar ~ sqft_living + region + waterfront + grade + yr_built, data=Data)
par(mfrow = c(2, 2))
plot(result2)
```
Simpler model:
```{r}
result3<-lm(ystar ~ sqft_living + region + waterfront, data=Data)
par(mfrow = c(2, 2))
plot(result3)
```
To do:
train/test data split
Another variable for evaluating fit?
Another method for automatic model fit?
automatic after transforming
add ##correlations between quantitative predictors
round(cor(train[,5:8], use= "complete.obs"),3)
