---
title: "Project2_Section4"
author: "Margaret Lindsay"
date: "2024-11-24"
output: html_document
---
## Loading Data/Packages
```{r}
library(tidyverse)
library(ggplot2)

Data <- read.csv("kc_house_data.csv",header = TRUE)
zipcode <- read.csv("zipcodeData.csv")
```

## Adding region predictor
```{r}
Data$region <- c()
for (i in 1:length(Data$zipcode)){
  regionMatch <- zipcode[zipcode$x == Data$zipcode[i],]$region
  Data$region[i]<- regionMatch 
}
```

## Adding Variables for Grade Categories
Grade is an index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design.
```{r}
Data <- Data %>% 
  mutate(grade.cat=case_when(grade <= 3 ~ "low",
    grade >3 & grade <=6 ~ "medium low", grade == 7 ~ "average", grade > 7 & grade < 11 ~ "medium high", grade >= 11 ~ "high"))

Data$grade.cat<-factor(Data$grade.cat, levels=c("low","medium low","average","medium high", "high"))
```

## Removing errors identified in section 3
```{r}
cleanData <- Data[(Data$bedrooms!=0),]
cleanData <- cleanData[(cleanData$bathrooms!=0),]
cleanData <- cleanData[(cleanData$bedrooms<30),]
cleanData <- cleanData[ , -which(names(cleanData) %in% c("date","sqft_above","sqft_basement","zipcode","lat","long"))]
```

## Split for Training/Test Data
```{r}
set.seed(6021)
sample.data<-sample.int(nrow(cleanData), floor(.50*nrow(cleanData)), replace = F)
train<-cleanData[sample.data, ]
test<-cleanData[-sample.data, ]
```

## Univariate Visualizations
```{r}
summary(train$price)

ggplot(train, aes(y=price))+
  geom_boxplot()+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_y_continuous(labels = scales::comma)+
  labs(y="Price ($)\n", title="Box Plot of Selling Price")
```

```{r}
ggplot(train,aes(x=price)) + 
  geom_density()+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_continuous(labels = scales::comma)+
  labs(x="Price ($)", y="Density", title="Density Plot of Selling Price")
```

```{r}
ggplot(train,aes(x=price))+
  geom_histogram(binwidth = 10000, color='navy')+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_continuous(labels = scales::comma)+
  labs(x="Price ($)", y="Count", title="Histogram of Selling Price")
```

```{r}
train.under2M <- train %>% 
  filter(price < 2000000) %>% 
  arrange(-price)

ggplot(train.under2M,aes(x=price))+
  geom_histogram(binwidth = 10000, color='navy')+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_continuous(labels = scales::comma)+
  labs(x="Price ($)", y="Count", title="Histogram of Selling Price")
```


```{r}
ggplot(train,aes(x=yr_built)) + 
  geom_density()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Year Built", y="Density", title="Density Plot of House Counstruction Year")
```

```{r}
ggplot(train,aes(x=yr_built))+
  geom_histogram(binwidth = 10, color='white')+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Year Built", y="Count", title="Histogram of House Counstruction Year")
```

```{r}
ggplot(train,aes(x=bedrooms))+
  geom_histogram(binwidth=1, color='white')+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Number of Bedrooms", y="Count", title="Histogram of Number of Bedrooms")+
  stat_count(binwidth = 1, 
             geom = 'text', 
             color = 'white', 
             aes(label = after_stat(count)),
            position = position_stack(vjust = 0.5))

```

## Bivariate Visuaizations
```{r}
ggplot(train, aes(x=bedrooms, y=price))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Number of Bedrooms", y="Price ($)\n", title="Selling Price Against Number of Bedrooms")
```

```{r}
ggplot(train, aes(x=bathrooms, y=price))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Number of Bathrooms", y="Price ($)\n", title="Selling Price Against Number of Bathrooms")
```

```{r}
ggplot(train, aes(x=sqft_living, y=price))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  scale_x_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Interior Square Footage", y="Price ($)\n", title="Selling Price Against Interior Square Footage")

Data %>% 
  arrange(-sqft_living)
```

```{r}
ggplot(train, aes(x=grade, y=price))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Grade", y="Price ($)\n", title="Home Selling Price Against Grade")
```

```{r}
train$waterfront <- factor(train$waterfront)

ggplot(train, aes(x=waterfront, y=price))+
  geom_boxplot()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Waterfront Status", y="Price ($)\n", title="Home Selling Price Against Waterfront Status")
```
```{r}
train$view <- factor(train$view)

ggplot(train, aes(x=view, y=price))+
  geom_boxplot()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="View Ranking", y="Price ($)\n", title="Home Selling Price Against View Ranking")
```


```{r}
ggplot(train, aes(x=grade.cat, y=price))+
  geom_boxplot()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Grade Category", y="Price ($)\n", title="Home Selling Price Against Grade")
```


```{r}
ggplot(train, aes(x=condition, y=price))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Condition", y="Price ($)\n", title="Home Selling Price Against Home Condition")
```


```{r}
train$condition <- factor(train$condition)
ggplot(Data, aes(x=condition, y=price))+
  geom_boxplot()+
  scale_y_continuous(labels = scales::comma, limits=c(0,2000000))+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Condition", y="Price ($)\n", title="Home Selling Price Against Home Condition")
```
```{r}
#tapply(Data$price,Data$condition,mean)
tapply(train$price,train$condition,median)
```

## Multivariate Visualizations
```{r}
train <- train %>% 
  #mutate(ouMean = ifelse(price>mean(price),"Over Mean","Under Mean"))
  mutate(ouMedian = ifelse(price>median(price),"Over Median","Under Median"))
  
ggplot(train,aes(x=yr_built, color=ouMedian)) + 
  geom_density()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Year Built", y="Density", title="Density Plot of House Counstruction Year", color="Relation to Median Sale Price")
```
```{r}
ggplot(train, aes(x=bedrooms, y=price, color=grade))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Number of Bedrooms", y="Price ($)\n", color="Grade", title="Home Selling Price Against Number of Bedrooms by Grade")
```

```{r}
ggplot(train, aes(x=sqft_lot, y=price, color=grade))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  scale_x_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Lot Square Footage", y="Price ($)\n", color="Grade", title="Home Selling Price Against Lot Sqaure Footage by Grade")
```

```{r}
ggplot(train, aes(x=bedrooms, y=price, color=condition))+
  geom_point(alpha=0.5)+
  scale_y_continuous(labels = scales::comma)+
  scale_x_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Number of Bedrooms", y="Price ($)\n", color="Condition", title="Home Selling Price Against Number of Bedrooms by Condition")
```

```{r}
ggplot(train, aes(x=sqft_living, y=price, fill=bedrooms, size=condition))+
  geom_point(shape=21,alpha=0.5)+
  scale_y_continuous(labels = scales::comma)+
  scale_x_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Interior Square Footage", y="Price ($)\n", color="Bedrooms", title="Selling Price Against Number of Bedrooms by Condition")
```

