---
title: "Project2_Sections4_7"
author: "Margaret Lindsay"
date: "2024-11-30"
output: html_document
---
# Loading and cleaning data

```{r}
## Loading Data/Packages
library(tidyverse)
library(ggplot2)

Data <- read.csv("kc_house_data.csv",header = TRUE)
zipcode <- read.csv("zipcodeData.csv")

## Adding region predictor
Data$region <- c()
for (i in 1:length(Data$zipcode)){
  regionMatch <- zipcode[zipcode$x == Data$zipcode[i],]$region
  Data$region[i]<- regionMatch 
}

## Adding Binary Variable for Good Quality Homes (Condition > 3 & Grade > 7)
Data <- Data %>% 
  mutate(goodQuality = ifelse(condition > 3 & grade > 7, "Yes", "No"))

## Adding Variables for Grade Categories
Data <- Data %>% 
  mutate(grade.cat=case_when(grade <= 3 ~ "low",
    grade >3 & grade <=6 ~ "medium low", grade == 7 ~ "average", grade > 7 & grade < 11 ~ "medium high", grade >= 11 ~ "high"))

Data$grade.cat<-factor(Data$grade.cat, levels=c("low","medium low","average","medium high", "high"))

Data$goodQuality <- factor(Data$goodQuality)

## Factoring Categorical Variables
Data$waterfront <- factor(Data$waterfront)

## Removing errors identified in section 3
cleanData <- Data[(Data$bedrooms!=0),]
cleanData <- cleanData[(cleanData$bathrooms!=0),]
cleanData <- cleanData[(cleanData$bedrooms<30),]
cleanData <- cleanData[ , -which(names(cleanData) %in% c("date","sqft_above","sqft_basement","zipcode","lat","long"))]

## Split for Training/Test Data

set.seed(6021)
sample.data<-sample.int(nrow(cleanData), floor(.50*nrow(cleanData)), replace = F)
train<-cleanData[sample.data, ]
test<-cleanData[-sample.data, ]

## Unbalanced? -- Yes, unbalanced toward No
table(cleanData$goodQuality)
```


# SECTION 4

## Univariate Visualizations
```{r}
summary(train$price)

ggplot(train, aes(y=price))+
  geom_boxplot()+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_y_continuous(labels = scales::comma)+
  labs(y="Price ($)\n", title="Box Plot of Selling Price")
```

```{r}
ggplot(train,aes(x=price)) + 
  geom_density()+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_continuous(labels = scales::comma)+
  labs(x="Price ($)", y="Density", title="Density Plot of Selling Price")
```

```{r}
ggplot(train,aes(x=price))+
  geom_histogram(binwidth = 10000, color='navy')+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_continuous(labels = scales::comma)+
  labs(x="Price ($)", y="Count", title="Histogram of Selling Price")
```

```{r}
train.under2M <- train %>% 
  filter(price < 2000000) %>% 
  arrange(-price)

ggplot(train.under2M,aes(x=price))+
  geom_histogram(binwidth = 10000, color='navy')+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_continuous(labels = scales::comma)+
  labs(x="Price ($)", y="Count", title="Histogram of Selling Price")
```

```{r}
ggplot(train,aes(x=yr_built)) + 
  geom_density()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Year Built", y="Density", title="Density Plot of House Counstruction Year")
```

```{r}
ggplot(train,aes(x=yr_built))+
  geom_histogram(binwidth = 10, color='white')+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Year Built", y="Count", title="Histogram of House Counstruction Year")
```

```{r}
ggplot(train,aes(x=bedrooms))+
  geom_histogram(binwidth=1, color='white')+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Number of Bedrooms", y="Count", title="Histogram of Number of Bedrooms")+
  stat_count(binwidth = 1, 
             geom = 'text', 
             color = 'white', 
             aes(label = after_stat(count)),
            position = position_stack(vjust = 0.5))
```

## Bivariate Visuaizations
```{r}
ggplot(train, aes(x=bedrooms, y=price))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Number of Bedrooms", y="Price ($)\n", title="Selling Price Against Number of Bedrooms")
```

```{r}
ggplot(train, aes(x=bathrooms, y=price))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Number of Bathrooms", y="Price ($)\n", title="Selling Price Against Number of Bathrooms")
```

```{r}
ggplot(train, aes(x=sqft_living, y=price))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  scale_x_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Interior Square Footage", y="Price ($)\n", title="Selling Price Against Interior Square Footage")

Data %>% 
  arrange(-sqft_living)
```

```{r}
ggplot(train, aes(x=grade, y=price))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Grade", y="Price ($)\n", title="Home Selling Price Against Grade")
```

```{r}
ggplot(train, aes(x=waterfront, y=price))+
  geom_boxplot()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Waterfront Status", y="Price ($)\n", title="Home Selling Price Against Waterfront Status")
```
```{r}
ggplot(train, aes(x=factor(train$view), y=price))+
  geom_boxplot()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="View Ranking", y="Price ($)\n", title="Home Selling Price Against View Ranking")
```


```{r}
ggplot(train, aes(x=grade.cat, y=price))+
  geom_boxplot()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Grade Category", y="Price ($)\n", title="Home Selling Price Against Grade")
```


```{r}
ggplot(train, aes(x=condition, y=price))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Condition", y="Price ($)\n", title="Home Selling Price Against Home Condition")
```


```{r}
tapply(train$price,train$condition,median)
```

## Multivariate Visualizations
```{r}
train <- train %>% 
  mutate(ouMedian = ifelse(price>median(price),"Over Median","Under Median"))
  
ggplot(train,aes(x=yr_built, color=ouMedian)) + 
  geom_density()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Year Built", y="Density", title="Density Plot of House Counstruction Year", color="Relation to Median Sale Price")
```
```{r}
ggplot(train, aes(x=bedrooms, y=price, color=grade))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Number of Bedrooms", y="Price ($)\n", color="Grade", title="Home Selling Price Against Number of Bedrooms by Grade")
```

```{r}
ggplot(train, aes(x=sqft_lot, y=price, color=grade))+
  geom_point()+
  scale_y_continuous(labels = scales::comma)+
  scale_x_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Lot Square Footage", y="Price ($)\n", color="Grade", title="Home Selling Price Against Lot Sqaure Footage by Grade")
```

```{r}
ggplot(train, aes(x=bedrooms, y=price, color=condition))+
  geom_point(alpha=0.5)+
  scale_y_continuous(labels = scales::comma)+
  scale_x_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Number of Bedrooms", y="Price ($)\n", color="Condition", title="Home Selling Price Against Number of Bedrooms by Condition")
```

```{r}
ggplot(train, aes(x=sqft_living, y=price, fill=bedrooms, size=condition))+
  geom_point(shape=21,alpha=0.5)+
  scale_y_continuous(labels = scales::comma)+
  scale_x_continuous(labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x="Interior Square Footage", y="Price ($)\n", color="Bedrooms", title="Selling Price Against Number of Bedrooms by Condition")
```


#SECTION 7


## Fit Logistic Regression
```{r}
result<-glm(goodQuality ~ price + bedrooms + bathrooms + sqft_living + sqft_lot + floors + waterfront + view + yr_built + yr_renovated + sqft_living15 + sqft_lot15 + region, family=binomial, data=train)
summary(result)

## Check for multicolinearity-- evidence for multicolinearity (VIFs > 5)
library(faraway) 
faraway::vif(result)
```

## Reduced model

H0: coefficients for all dropped predictors = 0
Ha: at least one coefficient is not 0

Dropped: sqft_living, sqft_lot, waterfront, sqft_lot15 (4) since not significant

Since p-value is greater than 0.05, we cannot reject the null and therefore we can use the reduced model.
```{r}
reduced <- glm(goodQuality ~ price + bedrooms + bathrooms + floors + view + yr_built + yr_renovated + sqft_living15 + region, family=binomial, data=train)
summary(reduced)

faraway::vif(reduced)

test.stat <- reduced$deviance - result$deviance
p.val <- 1 - pchisq(test.stat, 4)
p.val
```

## Confusion Matrix
Data is unbalanced. Using a threshold of 0.5, the FNR was high, so we lowered the threshold.
```{r}
##predicted probs for test data
preds <- predict(reduced, newdata=test, type="response")

##confusion matrix with threshold of 0.5
table(test$goodQuality, preds>0.5)

##confusion matrix with threshold of 0.15
table(test$goodQuality, preds>0.15)
```

## ROC Curve
```{r}
library(ROCR)
rates <- ROCR::prediction(preds, test$goodQuality)
roc_result <- ROCR::performance(rates,measure="tpr", x.measure="fpr")
plot(roc_result, main="ROC Curve for Reduced Model")
lines(x = c(0,1), y = c(0,1), col="red")
```

## AUC
```{r}
auc <- ROCR::performance(rates, measure = "auc") 
auc@y.values
```


