---
title: "Group 2, Project 2 WIP"
author: "Nathan Koh"
date: "2024-11-17"
output: html_document
---
help


loading csv
```{r}
data<-read.csv("kc_house_data.csv", header=TRUE)
```

From Kaggle, variable definitions:

*id* - Unique ID for each home sold
*date* - Date of the home sale
*price* - Price of each home sold
*bedrooms* - Number of bedrooms
*bathrooms* - Number of bathrooms, where .5 accounts for a room with a toilet but no shower
*sqft_living* - Square footage of the apartments interior living space
*sqft_lot* - Square footage of the land space
*floors* - Number of floors
*waterfront* - A dummy variable for whether the apartment was overlooking the waterfront or not
*view* - An index from 0 to 4 of how good the view of the property was
*condition* - An index from 1 to 5 on the condition of the apartment,
*grade* - An index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design.
*sqft_above* - The square footage of the interior housing space that is above ground level
*sqft_basement* - The square footage of the interior housing space that is below ground level
*yr_built* - The year the house was initially built
*yr_renovated* - The year of the house’s last renovation
*zipcode* - What zipcode area the house is in
*lat* - Latitude
*long* - Longitude
*sqft_living15* - The square footage of interior housing living space for the nearest 15 neighbors
*sqft_lot15* - The square footage of the land lots of the nearest 15 neighbors


Observations WIP:

1) The Bathroom column indicates 0.5 represents a bathroom with a toilet but no shower. Data shows 0.25 increments.

2) Comparing price to other columns may not make sense

3) The column Bedroom has a similar problem to Bathroom

3.5) Bedroom, Bathroom, and the Price on certain homes do not make sense

4) Floors indicate a 0.5th floor but is not described as to what it represents


Possible solutions:

1) Can't delete them since Obs 1 has ~7000 results using only .25 increments


Removing the following vars: yr_built, yr_renovated, zipcode,lat, long. Location and calendar data is not as important
```{r}
library(dplyr)

data<- data %>% select(-yr_built,-yr_renovated,-zipcode,-lat,-long)

data
```

