---
title: "Project2_Section7"
author: "Margaret Lindsay"
date: "2024-11-30"
output: html_document
---


```{r}
## Loading Data/Packages
library(tidyverse)
library(ggplot2)

Data <- read.csv("kc_house_data.csv",header = TRUE)
zipcode <- read.csv("zipcodeData.csv")

## Adding region predictor
Data$region <- c()
for (i in 1:length(Data$zipcode)){
  regionMatch <- zipcode[zipcode$x == Data$zipcode[i],]$region
  Data$region[i]<- regionMatch 
}

## Adding Binary Variable for Good Quality Homes (Condition > 3 & Grade > 7)
Data <- Data %>% 
  mutate(goodQuality = ifelse(condition > 3 & grade > 7, "Yes", "No"))

Data$goodQuality <- factor(Data$goodQuality)
```

## Factoring Categorical Variables

```{r}
Data$waterfront <- factor(Data$waterfront)
```

## Removing errors identified in section 3

```{r}
cleanData <- Data[(Data$bedrooms!=0),]
cleanData <- cleanData[(cleanData$bathrooms!=0),]
cleanData <- cleanData[(cleanData$bedrooms<30),]
cleanData <- cleanData[ , -which(names(cleanData) %in% c("date","sqft_above","sqft_basement","zipcode","lat","long"))]
```

## Split for Training/Test Data

```{r}
set.seed(6021)
sample.data<-sample.int(nrow(cleanData), floor(.50*nrow(cleanData)), replace = F)
train<-cleanData[sample.data, ]
test<-cleanData[-sample.data, ]
```

## Unbalanced?

Unbalanced toward No?
```{r}
table(cleanData$goodQuality)
```

## Fit Logistic Regression

```{r}
result<-glm(goodQuality ~ price + bedrooms + bathrooms + sqft_living + sqft_lot + floors + waterfront + view + yr_built + yr_renovated + sqft_living15 + sqft_lot15 + region, family=binomial, data=train)
summary(result)
```

## Check for multicolinearity

Lots of evidence for multicolinearity (VIFs > 5)
```{r}
library(faraway) 
faraway::vif(result)
```

## Reduced model

H0: coefficients for all dropped predictors = 0
Ha: at least one coefficient is not 0

Dropped: sqft_living, sqft_lot, waterfront, sqft_lot15 (4) since not significant

Since p-value is greater than 0.05, we cannot reject the null and therefore we can use the reduced model.
```{r}
reduced <- glm(goodQuality ~ price + bedrooms + bathrooms + floors + view + yr_built + yr_renovated + sqft_living15 + region, family=binomial, data=train)
summary(reduced)

faraway::vif(reduced)

test.stat <- reduced$deviance - result$deviance
p.val <- 1 - pchisq(test.stat, 4)
p.val
```
# Keep sqft_lot??

## Confusion Matrix
Data is unbalanced. Using a threshold of 0.5, the FNR was high, so we lowered the threshold.
```{r}
##predicted probs for test data
preds <- predict(reduced, newdata=test, type="response")

##confusion matrix with threshold of 0.2
table(test$goodQuality, preds>0.15)
```

## ROC Curve

Looks pretty good--"The further the curve is from the diagonal and closer to the top left of the plot, the better the model is in classifying observations correctly."
```{r}
library(ROCR)
rates <- ROCR::prediction(preds, test$goodQuality)
roc_result <- ROCR::performance(rates,measure="tpr", x.measure="fpr")
plot(roc_result, main="ROC Curve for Reduced Model")
lines(x = c(0,1), y = c(0,1), col="red")
```

## AUC

"An AUC closer to 1 indicates the model does better than random guessing in classifying observations."
```{r}
auc <- ROCR::performance(rates, measure = "auc") 
auc@y.values
```


